---
title: "Statistical Inferences and simulations"
author: "Shaun"
date: "February 19, 2017"
output: html_document
---
##Overview
For the purpose of this assignment, we'll be exploring the estimation of population parameters from samples by using the Central Limit Theorem. It will be shown that, given enough samples, the mean of will be approximately normal and act as a good estimate for the population.

##Simulations
The following code generates a dataframe with a 1000 rows of 40 randomly generated numbers from an exponential distribution.
It then calculates the mean and standard deviation of each row.

```{r echo=TRUE}
# Load required libraries
library(ggplot2)
library(gridExtra)
library(boot)
n <- 40             # Number of exponentials to generate per simulation
lambda <- 0.2       # Rate parameter for rexp function
mu <- 1/lambda      # mean 
sig <- 1/lambda     # standard deviation
vars <- sig^2       # calculate the variance
sim.count <- 1000   # Number of simulation to run
dis.matrix <- NULL  # Initialising data frame

# Run simulation to generate a sim.count by n (1000 x 40) data frame 
for(i in 1:sim.count){
  dis.matrix <- rbind(dis.matrix,rexp(n, lambda))
}
# Calculate the row means
rms <- as.data.frame(rowMeans(dis.matrix))
colnames(rms) <- "mean"

# Calculate the row standard deviations
sds <- as.data.frame(apply(dis.matrix, 1, sd))
colnames(sds) <- "std"
stack()
t.test(rms$mean)$conf.int
```

## Graphs

```{r echo=TRUE}
mean.plot <- ggplot(rms, aes(x= rms))+
  geom_histogram(fill="blue", colour="black", bins = 30)+
  scale_x_continuous()+
  labs(title="Histogram of mean distribution", x="Mean", y="Count")+
  geom_vline(xintercept = mu, colour="red")+
  geom_vline(xintercept = mean(as.matrix(rms)), colour="green")+
  geom_text(aes(x=mu, label=paste0("Theoretical mean = ", mu), y=10), colour="red")+
  geom_text(aes(x=mean(as.matrix(rms)), label=paste0("Sample mean = ", round(mean(as.matrix(rms)),2)), y=30), colour="green")

exp.plot <- ggplot(as.data.frame(dis.matrix), aes(x= as.data.frame(dis.matrix)))+
  geom_histogram(fill="blue", colour="black", bins = 30)+
  scale_x_continuous()+
  labs(title="Histogram of exp distribution", x="Exponential value", y="Count")

grid.arrange(exp.plot, mean.plot)

```


```{r echo=TRUE}
ggplot(sds, aes(x = sds^2))+
  geom_histogram(fill="blue", colour="black", bins=30)+
  scale_x_continuous()+
  labs(title="Histogram of the variance distribution", x = "Variance", y = "Count")+
  geom_vline(xintercept = vars, colour = "red")+
  geom_vline(xintercept = mean(sds^2), colour="green")+
  geom_text(aes(x = vars, label=paste0("Theoretical variance = ", vars), y=10), colour="red")+
  geom_text(aes(x = mean(sds^2), label=paste0("Sample variance = ", round(mean(sds^2),2)), y=30), colour="green")
```